---
title: "Investigating adding postcodes to maps"
author: "Laura Flenley"
date: "2024-12-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE)
#include = TRUE: means that any outputs or graphs produced from the code will be shown in the final document
#echo = FALSE: means that the code used to produce the outputs will not be printed in the final document 

#chunk summary: in general throughout the document, code outputs will be shown, but not the code used to produce them: unless specified otherwise in a specific chunk
```

<!--### 1. Install packages--> 

```{r install packages, eval = FALSE}
#code for installing packages
install.packages("tidyverse")
install.packages("sf")

#this code will not run unless you change 'eval = TRUE' in the chunk setup: because I already have the packages installed on my computer, so I do not need to install them each time I run the code. They are included so that anyone else looking to run this code can install the packages if they do not have them installed already.

#Packages are collections of code instructions that someone else has helpfully made, that make it easier to do particular things - like a shortcut. I will use the 'tidyverse' package (has code for plotting graphs) and 'sf' package (has code for making maps alongside the tidyverse package).

#Packages need to be 'installed' (aka downloaded) onto a computer once, but not a second time. 

```

<!-- ### 2.  Load packages and data -->

**Dataset / Map data:** I have downloaded some datasets: 

- **'UK postcode boundary polygons'** from the **Open Door Logistics** website, which is described as 'reconstructed UK postcode polygons (January 2015) in Shapefile format'. There may be some terms and conditions to using this data that we need to be aware of.

- **'Indices of Multiple Deprivation'** from the **UK Government via Communities Open Data**, as part of the ArcGis Community Hub. This contains the indices of multiple deprivation as shapefiles.

- **'Code-point open'** from the **Ordinance Survey Data Hub**, which contains extra details of postcodes, not as shapefiles but as a csv datatable. 

```{r load packages & data, subset data include=FALSE}
#load packages
library(tidyverse)
library(sf)
library(dplyr)
library(stringr)

#load the data into R, and name it
postcodes_sf<-read_sf("data/postcode_boundary_data/Distribution/Sectors.shp") #this is postcode sector info
multdep<-read_sf("data/multdep_2019/Indices_of_Multiple_Deprivation_(IMD)_2019/Indices_of_Multiple_Deprivation_(IMD)_2019.shp")

#have a look at it 
head(postcodes_sf)
head(multdep)

#subset the postcodes data for oxfordshire:
  ## Use dplyr and stringr to filter the postcodes data for rows with oxford postcodes: aka observations with 'OX' in them
oxford_postcodes_sf <- postcodes_sf%>%
  filter(str_detect(name, regex("OX", ignore_case = TRUE)))
#check it worked
unique(oxford_postcodes_sf)

#subset the IMD data for oxfordshire:
oxonly_multdep<-multdep %>%
  filter(str_detect(LADnm, regex("oxford|vale of white horse|cherwell", ignore_case = TRUE)))
#check it worked
unique(oxonly_multdep$LADnm)

#subset the IMD data for a larger subset (WIP):
ox_multdep<-multdep %>%
  filter(str_detect(LADnm, regex("oxford|vale of white horse|cherwell|south buckingham|chiltern|wycombe|aylesbury vale|eastleigh|east hampshire|basingstoke and deane|fareham|gosport|^hart$|havant|New forest|portsmouth|rushmoor|southampton|test valley|winchester", ignore_case = TRUE)))
#check it worked
unique(ox_multdep$LADnm)

#testing testing 123

```

### Initial graphs using IMD data

```{r initial maps, eval=TRUE}

#plot for the strictly oxfordshire IMD area
plot <- ggplot(oxonly_multdep)+ 
  geom_sf(aes(fill = oxonly_multdep$IMD_Decile), colour = NA)+ #colour the IMD shapes based on IMD value, and don't give them a border
  scale_fill_continuous(name = "Index of Multiple Deprivation, 2019", type = "viridis", direction = -1)+ #IMD fill should be given a name; filled using the colour palette 'viridis', but so that high values are darker and low values are lighter: this is opposite to how it is usually applied, hence use direction = -1
  geom_sf(data=oxford_postcodes_sf, colour = "white", alpha=0.1)+ #add postcode sectors on top, in white outline
    theme_minimal()
print(plot)


```

### Focussing on Oxfordshire constituencies

(as an example of a smaller map - ideally would include the wider places people travel from rather than just Oxfordshire)
```{r map 2: oxfordshire, eval=TRUE}

#make oxfordshire dataset into a map
ggplot(oxfordshire_sf) + #using the oxfordshire dataset, make a plot
  geom_sf(fill = "#3BA740", colour = "white")+ #fill everything with green, and include white outlines to constituencies
  labs(title = "Constituencies in Oxfordshire")+ #specify title of graph
  theme_minimal() #give it a white background


```

### ...with some details added

- to show how points / locations / lines can be added: I've added my house & Dorchester Abbey & connected them.

- I've added the electorate of the different constituencies to show how you can colour-code different areas

- you can also overlay different maps together, which I haven't done here. For instance, I could presumably show more features of the constituencies than just the outlines

```{r oxfordshire map with detail, fig.height = 6*1.5, fig.width = 7*1.5}

#enter here some fancy coding to make the names of constituencies appear
oxfordshire_sf$centroid <- st_centroid(oxfordshire_sf$geometry) #add a column called 'centroid' to our oxfordshire dataset. this column should contain the information about the location of the center of each constituency. We're telling R to calculate these centre locations, using the readymade function st_centroid() and the 'geometry' column that is already in our dataset 

centroids <- st_coordinates(oxfordshire_sf$centroid) #make a new data column called centroids. we want it to contain the coordinates of the centroids, which it should obtain from the centroid column we just made before. R will work out the coordinates by using the function st_coordinates()
oxfordshire_sf <- cbind (oxfordshire_sf, centroids) #add the centroids data object we just made to the oxfordshire dataset: now we know the coordinates of the centre of each constituency, so that we can position the constituency names in the middle of each constituency.

#create a new dataset with an added column of names, so the longest names can have line breaks incorporated when we make our plot = looks nicer
oxfordshire_sf_named<-oxfordshire_sf %>% #new dataset is called oxfordshire_sf_named, which starts off as being a copy of oxfordshire_sf
  mutate(name = case_when( #change some parts of the new dataset, if:
    PCON24NM == "Bicester and Woodstock" ~ "Bicester and\nWoodstock",  #if the name is Bicester and Woodstock, change the name to Bicester and \nWoodstock
      #\n indicates a line break: so when we make the graph, the name will not be one very long thin line
    PCON24NM == "Henley and Thame" ~ "Henley and \nThame", #same for this
    PCON24NM == "Oxford West and Abingdon" ~ "Oxford West and\nAbingdon", #same for this
    TRUE ~ PCON24NM #for all the rest of the constituencies: keep the name the same (no linebreak - because I've decided they don't need one)
  ))

#adding dorchester abbey & laura's house
interest_points <- data.frame( #make a new dataframe called interest_points, which will have in it:
  Location = c("Dorchester Abbey", "Laura's house"), #it should have a column called location, with dorchester abbey & laura's house as rows
  lat = c(51.646301, 51.741770), #it should have a column called lat, with these values (latitude - found on google)
  lon = c(-1.167200, -1.258005) #it should have a column called lon, with these values (longitude - found on google)
)

#turn into a new sf object: I don't fully understand what this means, but R needs us to do this if we want to mak e the dataset into a map
interest_points_sf <- st_as_sf(interest_points, coords = c("lon", "lat"), crs = 4326)
  #crs = 4326 states these coordinates we've given it represent longitude & latitude on earth's surface

#convert interest points to same crs as base map otherwise all R-hell breaks loose apparently 
interest_points_sf <- st_transform(interest_points_sf, st_crs(oxfordshire_sf_named))
#interest_points_sf

#make a table of just the relevant coordinates of the interest points to add the points to the map
interest_points_coords <- st_coordinates(interest_points_sf) #new table has the coordinates we just put into our interest_points_sf file; but specifically just as numbers rather than some cryptic coding language thing
#interest_points_coords

#lets add electorate to the dataframe oxfordshire_sf_named for fun! yay
Electorate <- c(69943, 70389, 74356,70626,72371,72004,70042) #make a new table with the electorate, listed in the same order as the constituencies appear in the table (alphabetically). this data comes from wikipedia
oxfordshire_sf_named <- cbind(oxfordshire_sf_named, Electorate) #join on the electorate as a column to our master oxfordshire dataset

#define laura's journey distance as the crow flies
journey_line_laura <- st_sfc(st_linestring(interest_points_coords)) #tell it to make a line, using the coordinates of the two points, from our coordinates table
#reassure R that the line has the same CRS as the other datasets (because we already converted it to the right crs before extracting the coordinates)
journey_line_laura <- st_set_crs(journey_line_laura, st_crs(oxfordshire_sf_named))

```

Now we've done all the set up, we can code the plot:

```{r big lovely plot, fig.height = 6*1.5, fig.width = 7*1.5}

#big lovely plot
ggplot(oxfordshire_sf_named) + #make a plot, using data from oxfordshire_sf_named
  
  geom_sf(aes(fill = Electorate), color = "white")+ #colour fill the constituencies based on the electorate values, and give them white boundaries
  
  scale_fill_continuous(name = "Electorate", high = "#8788ab", low = "#cacbdf")+ #higher electorate values should be darker purple, lower values should be lighter purple
  
  geom_text(aes(x=X, y=Y, label = name), size = 3, colour = "black", nudge_y=+2500, nudge_x=-1500)+ #add the constituency names, using the 'name' column we added (has the line breaks we want). put them text size 3, black text, and move them a bit up & left (so they don't all overlap other text)
  
  geom_point(data = interest_points_sf, aes(x = interest_points_coords[,1], y = interest_points_coords[,2]), colour = "#770088", size = 3, pch = 19) + #add our points of interest, with size 3 purple dots, using the data from the table we made (interest_points_sf), and getting the coordinates from the coordinates table we made
  
  geom_label(data = interest_points_sf, aes(x=interest_points_coords[,1], y = interest_points_coords[,2], label = Location), size = 3, nudge_y = -1300, nudge_x = 3500, colour = "#770088", label.padding = unit(0.1, "lines"), label.size=0, alpha = 0.5)+ #add names to the locations: names should be the 'Location' column from the data table interest_points_sf, using the same coordinates as the points (from the coordinates data interest_points_coords). don't put lines round the outside of labels, make the labels a bit transparent, and move them down and right a bit so they don't overlap the actual points
  
  labs(title = "Laura's journey (as the crow flies)", x="", y="")+ #add a title, and take away the x and y axis names - otherwise it just says X and Y
  
  theme_minimal()+ #white background
  
  geom_sf(data=journey_line_laura, colour="red", size = 1) #add the line between the points: from the line dataset we specified; make it size 1, and red.


```

### Some thoughts

- the general gist is you can customise almost everything about them

- There's a lot more information online about this type of coding so I can look into other ways I could add to this that would be relevant to us

- I can also look for data with more relevant/detailed information than the constituencies dataset I've used here
 
- I don't like how the constituency names appear.. that's a work in progress! They're unreasonably fiddly to move around

<!-- **REMOVED GRAPHS** 

1. an attempt to move one label down a bit - worked but duplicated the name so it was on there twice.
ggplot(oxfordshire_sf_named) +
geom_sf(fill = "#3BA740", colour = "white")+
geom_text(aes(x=X, y=Y, label = name), size = 3, colour = "black")+
geom_text(data = oxfordshire_sf_named %>% filter(name == "Oxford West and\nAbingdon"),
aes(x = X, y = Y - 2500, label = name),  
size = 3, colour = "red")+
labs(title = "Constituencies in Oxfordshire")+
theme_minimal()

#trying to move it down - didn't work
ggplot(oxfordshire_sf_named) +
  geom_sf(fill = "#3BA740", colour = "white") +  # Base map
  geom_text(aes(x = X, y = Y, label = name), size = 3, colour = "black") +  # Original labels in black
  # Adjust label for "Bicester and Woodstock" using mutate or case_when in the aes
  geom_text(aes(x = X, y = ifelse(name == "Bicester and Woodstock", Y - 1, Y),  # Move down for "Bicester and Woodstock"
                label = name), 
            size = 3, colour = ifelse(oxfordshire_sf_named$name == "Bicester and Woodstock", "red", "black"))  # Change color conditionally

#trying to have the graph without the original label for oxford west and abingdon, so it could be added after. but it didn't work.
ggplot(oxfordshire_sf_named) +
  geom_sf(fill = "#3BA740", colour = "white")+
  geom_text(data = oxfordshire_sf_named %>% filter(name != "Oxford West and Abingdon"), aes(x=X, y=Y, label = name), size = 2, colour = "black")+
  labs(title = "Constituencies in Oxfordshire")+
  theme_minimal()
  
##attempt to change the coordinates of the centroid - didn't work - chatgpt was not referring to the centroid coordinate in the right way. we got closer but i'm fed up

oxfordshire_sf_named_adjusted <- oxfordshire_sf_named %>%
  mutate(Y_centroid = ifelse(name == "Oxford West and Abingdon", Y_centroid - 2500, Y_centroid)  # Adjust Y for Oxford West and Abingdon
  )

oxfordshire_sf_named_adjusted %>%
  filter(name == "Oxford West and\nAbingdon") %>%
  select(name, Y)

head(oxfordshire_sf_named)

ggplot(oxfordshire_sf_named_adjusted) +
  geom_sf(fill = "#3BA740", colour = "white")+
  geom_text(aes(x=X, y=Y, label = name), size = 2, colour = "black")+
  labs(title = "Constituencies in Oxfordshire")+
  theme_minimal()

-->



